### 2024-08-26 research/notes.tex
created report & exported as latex & pdf

% Created 2024-08-25 Sun 14:46
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Mark Krutzler}
\date{\today}
\title{Notes / Research on CPU-Scheduling}
\hypersetup{
 pdfauthor={Mark Krutzler},
 pdftitle={Notes / Research on CPU-Scheduling},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.4 (Org mode 9.8)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\part{Basics (OStep)}
\label{sec:orgce4bae2}
\chapter{Introduction (Chapter 7)}
\label{sec:org02640b6}
Scheduling is not a low level mechanism but a high level policy/disciplines.
We need to make simplifying assumptions of the workload:
\begin{enumerate}
\item Each job runs for the same amount of time
\item All jobs arrive at the same time.
\item Once started, each job runs to completion
\item All jobs only use the CPU (no I/O)
\item The run-time of each job is known.
\end{enumerate}
These rules will be eased over time until we get a fully functioning policy.
Of course the more you know the easier it is to schedule.
\section{Metrics}
\label{sec:orgb0f9911}
The fundamental question is: How to we measure the ``efficiency'' or the ``quality'' of the scheduler?
You can measure performance or fairness. Here are ways to measure performance:
\begin{itemize}
\item turnaround time
\begin{itemize}
\item Calculated as:
\(T_{turnaround} = T_{completion} - T_{arrival}\)
For us \(T_{arrival} = 0\), because of simplification 1. (can be neglected later)
\end{itemize}
\item response time: measures the frustration of the user, while looking at the spinning ball
\begin{itemize}
\item Calculated as:
\(T_{response} = T_{firstrun} - T_{arrival}\)
\begin{itemize}
\item For modern computers, it is essential that this is kept at a minimum
\end{itemize}
\end{itemize}
\item fairness: first job to finish divided by last job to finish (this is not a performance metric!!)
\end{itemize}
\section{First In, First Out (FIFO) / First Come, First Served (FCFS)}
\label{sec:org930f0a2}
\begin{itemize}
\item Most basic scheduling policy
\item Given our simplification it works really well and is easy to implement
\item However after relaxing assumption 1, it will perform poorly if a huge process gets infront of many small ones
\begin{itemize}
\item This is the so called \textbf{convoy effect}
\item It is like if you're at waiting in line to pay and before you have a family of five with two full carts: annoying
\end{itemize}
\end{itemize}
\section{Shortest Job First (SJF)}
\label{sec:org1f10514}
\begin{itemize}
\item The shortest job is run first
\begin{itemize}
\item non-preemptive: runs a process until finish
\item preemptive: can stop and perform a context switch
\end{itemize}
\item If the smaller tasks arrive later (by relaxing assumption 2), then we face the same problem as before. (due to this algorithm can't perform a context switch / is non-preemptive)
\end{itemize}
\section{Shortest Time-to-Completion (STCF) / Preemptive Shortest Job First (PSJF)}
\label{sec:orgb6c2621}
\begin{itemize}
\item This policy requires that rule 3 is ignored.
\item This is the preemptive version of SJF.
\item It updates, whenever a new job arrives or one is finished
\end{itemize}
\section{Round Robin (RR) / time-slicing}
\label{sec:orgc7555bf}
\begin{itemize}
\item this policy runs each job for a specified ``time slice'' / ``scheduling quantum'' (introducing a variable)
\item general technique is called ``amortization''.
\item The shorter the time slice, the more responsive the system, however context switching costs CPU time aswell, so you'll need to balance out
\item RR is one of the worst policies for turnaround time
\item It gives up performance for fairness
\end{itemize}
\section{Relaxing Assumptions 4 \& 5}
\label{sec:orgac13a29}
\begin{enumerate}
\item assumption 4
\label{sec:org61b9a39}
\begin{itemize}
\item If a job waits for I/O than it is in a state called ``blocked''
\item While a job is waiting for I/O, the CPU can be passed onto somebodye else: ``overlapping''
\end{itemize}
\item assumption 5
\label{sec:org64b9783}
\begin{itemize}
\item we usually have no idea how long a job will take
\item This actually breaks most of our policies, because they all rely on knowing the length of the job (except RR)
\item Solution: Multi-Level Feedback Queue (MLFQ) \(\Rightarrow\) See next Chapter
\end{itemize}
\end{enumerate}
\chapter{Multi-Level Feedback Queue (Chapter 8)}
\label{sec:org36243f5}
\begin{itemize}
\item One of the most known Policies (Turning Awarded)
\item It tries to:
\begin{itemize}
\item optimize turn around time (without knowing the length of the job)
\item minimize response time
\end{itemize}
\end{itemize}
\section{(Basic) Rules of MLFQ}
\label{sec:orga3d0fb4}
\begin{itemize}
\item There are multiple queues and each has their priority level. (higher priority is preferred when switching)
\item If multiple jobs are on the same priority than RR (Round Robin) is used
\item Priorities can change over time.
\item Assume that if a job is resource intensive than it will stay as such. (The history of the job determines the future)
\end{itemize}
\section{Changing Priority}
\label{sec:org4946c2d}
\begin{itemize}
\item Depending on the CPU time usage, the priority changes
\item ``allotment'': time that a job can spend at a given priority before demotion.
\end{itemize}
\section{Priority Boost}
\label{sec:orgfe8a82e}
\begin{itemize}
\item to counter starvation of longer jobs every now and then all of the jobs are put into the priority queue
\item also this counters the fact that some programs might start non interactively and than turn into interactive (you know what I mean)
\end{itemize}
\section{``Better Accounting'' (Anti gaming)}
\label{sec:orgd12a267}
\begin{itemize}
\item to prevent people from abusing the allotment method and game the CPU, we need to update rule 4:
\item previous:
\begin{enumerate}
\item If a job uses up its allotment while running, its priority is reduced
\item If a job gives up the CPU before the allotment is up, it stays at the same priority
\end{enumerate}
\item new:
Once a job uses up its time allotment at a given level, its priority is reduced
\end{itemize}
\section{Summary of Rules (Copied out of the book)}
\label{sec:orga6a11c6}
\begin{enumerate}
\item If Priority (A) > Priority (B) \(\Rightarrow\) A runs \& B doesn't
\item If Priority (A) = Priority (B) \(\Rightarrow\) A \& B run in RR
\item When a job enters the system, it is placed at the highest priority
\item Once a job uses up its time allotment at a given level, its priority is reduced
\item After some period S, move all the jobs in the system to the topmost queue
\end{enumerate}
\section{Voo-Doo Constants}
\label{sec:org578a8d8}
These constants heavily change how effective the MLFQ is:
\begin{itemize}
\item scheduling quantum (RR)
\item amount of queues
\item when to priority boost
\item allotment (could change in every priority queue)
\end{itemize}
\chapter{Proportional Share (Chapter 9)}
\label{sec:org0922768}
\begin{itemize}
\item This is a fair scheduler
\begin{itemize}
\item The more/longer jobs run the fairer it becomes
\end{itemize}
\item literally just hold a lottery to determine which programs runs next
\item ``tickets'' represent the share of a resource that a process should recieve = it is like a currency
\begin{itemize}
\item the more tickets you hold, the higher the chance that you have a winning one
\item every time slice a new ticket is picked out as the winning ticket
\item more generally tickets can represent the share of something.
\end{itemize}
\item the tickets are handed out to the user, who than can allocate among their jobs
\begin{itemize}
\item the user can use their ``own'' tickets which will be converted into the global currency
\end{itemize}
\item ticket transfer can be used to boost a process
\begin{itemize}
\item think server / client => client give server their tickets, so that the server has a higher global share
\end{itemize}
\item ?? in a trusted environment you could also inflate your own tickets to boost you own CPU time
\end{itemize}
\section{Advantages of using randomness}
\label{sec:org26fc12b}
\begin{itemize}
\item no strange corner-case behaviors
\item lightweight
\item if the randomizing algorithm is quick than the speed is quick
\begin{itemize}
\item faster algorithms tend to be more like pseudo-random
\end{itemize}
\end{itemize}
\section{Implementation}
\label{sec:org36bc201}
\begin{enumerate}
\item requirements:
\label{sec:orgf067f7d}
\begin{itemize}
\item random number generator
\item data structure (to track the processes of the system)
\item amount of total number of tickets
\end{itemize}
\item sample code (copied):
\label{sec:orgc4e4c39}
\begin{verbatim}
// counter: used to track if we’ve found the winner yet
int counter = 0;

// winner: call some random number generator to
//         get a value >= 0 and <= (totaltickets - 1)
int winner = getrandom(0, totaltickets);

// current: use this to walk through the list of jobs
node_t *current = head;
while (current) {
    counter = counter + current->tickets;
    if (counter > winner)
        break; // found the winner
    current = current->next;
}
// ’current’ is the winner: schedule it...
\end{verbatim}
\end{enumerate}
\section{Assigning tickets}
\label{sec:orga879a75}
\begin{itemize}
\item Remains open for now
\end{itemize}
\section{Stride Scheduling}
\label{sec:org3130976}
\begin{itemize}
\item it is a deterministic fair-share scheduler
\begin{itemize}
\item while lottery scheduling achieves the proportions with probability (can be off), stride scheduling gets it right each time.
\item PROBLEM: you can't have a new job entering, because it will monopolize the CPU (due to low pass value)
\end{itemize}
\item bit tricky to understand: there is another article about it that I'll later read (under heading Lottery Scheduling)
\item how it works:
\begin{itemize}
\item each process has a stride to begin with (the more tickets the smaller the stride)
\item each time the process runs, it's counter (called ``pass'') get incremented by the value of the stride
\begin{itemize}
\item this is tracking its global progress
\end{itemize}
\item scheduler schedules according to the pass and the stride
\begin{itemize}
\item pick the lowest pass
\end{itemize}
\end{itemize}
\end{itemize}
\begin{enumerate}
\item pseudo-implementation (code copied)
\label{sec:org4c549fe}
\begin{verbatim}
current = remove_min(queue); // pick client with min pass
schedule(current); // run for quantum
current->pass += current->stride; // update pass using stride
insert(queue, current); // return current to queue
\end{verbatim}
\end{enumerate}
\section{Sidequest: Linux Completely Fair Scheduler (CFS)}
\label{sec:orgf8d6ba7}
\begin{itemize}
\item will talk about it later as well
\item every process has a counter called ``vruntime''
\begin{itemize}
\item as they run it increases
\item the process with the lowest ``vruntime'' is next
\begin{itemize}
\item PROBLEM: while waiting / in I/O the process vruntime is not increased: after coming back alive, it'll monopolize the CPU
\item SOLUTION: Once a process wakes up, it will take the lowest amount of vruntime
\item PROBLEM: short sleep will make it less fair for you for you
\end{itemize}
\end{itemize}
\item the switching is controlled through parameters:
\begin{itemize}
\item sched\textsubscript{latency}: dynamic time slice (is calculated), typically 48ms divided by n number of processes
\item sched\textsubscript{latency} basicly determines the maximal time frame until each process has run atleast once (if not controlled for minimun time slice)
\end{itemize}
\item There is also a minimal time slice:
\begin{itemize}
\item min\textsubscript{granularity} (set to usually 6ms) ensures that each process runs atleast a certain amount of time switching
\begin{itemize}
\item else the context switch would be too expensive
\item with this the scheduler becomes less fair, when only looking at the sched\textsubscript{latency}, however it is a good tradeoff
\end{itemize}
\end{itemize}
\item CFS utilizes the periodic timer interrupt. This means every 1ms it can wake up and determine what to do next
\end{itemize}
\begin{enumerate}
\item Niceness (Priority setting)
\label{sec:orgb6d68eb}
\begin{itemize}
\item priority setting is done through the ``nice'' level
\begin{itemize}
\item default: 0 (min: +19, max: -20)
\item the level will be mapped to a ``weight'' according to a premade table
\begin{itemize}
\item this will keep the proportianility
meaning: if you have a difference of 5 levels between two jobs, than the ratio of sharing stays the same
\end{itemize}
\item The time slice is calculated as followed:
$$timeslice_k = \frac{weight_k}{\sum_{i=0}^{n-1} weight_i} * schedlatency$$
\begin{itemize}
\item here n is the amount of processes
\end{itemize}
\item new vruntime is also calculated according to the niceness:
$$vruntime_i = vruntime_i + \frac{weight_0}{weight_i}*runtime_i$$
\end{itemize}
\end{itemize}
\item Efficiency of CFS (Red-Black Trees)
\label{sec:orgfa1c167}
\begin{itemize}
\item a scheduler has to make decisions as quickly as possible (this should hopefully be scaleable)
\item only runnable processes are kept here
(removed while waiting for I/O)
\item efficiency should be logarithmic (what does that mean?)
\item how does it even work?
\end{itemize}
\end{enumerate}
\part{Computer Scheduling Methods and their Countermeasures}
\label{sec:org853bfc0}
\begin{itemize}
\item Started reading it and it didn't really say anything new. I than scanned over it and skipped the rest
\end{itemize}
\chapter{Classification of Policies}
\label{sec:org3ff5e09}
\section{Characteristics}
\label{sec:org0b5c812}
\begin{itemize}
\item preemptive vs non-preemptive (already mentionned above)
\begin{itemize}
\item preemptive: if a higher priority exists, than the task can and will be abrupted
\item non-preemptive: opposite of preemptive
\end{itemize}
\item resume vs restart
\begin{itemize}
\item if a preempted job ``comes into service again'', should we resume where we left off or should we restart the whole thing?
\end{itemize}
\item where does priority come from?
\begin{itemize}
\item job environment (e.g.: running time, I/O)
\item computer system enviroment (dynamic priorities: e.g.: amount of jobs)
\item users environment (assigned by user)
\end{itemize}
\item knowledge of estimated time until finished
\begin{itemize}
\item most of the computers processes don't have a preset time
\end{itemize}
\end{itemize}
\section{Priority Based on running time only}
\label{sec:org7d8be45}
\begin{itemize}
\item gives shorter jobs and advantage
\end{itemize}
\begin{enumerate}
\item Shortest Job First (SJF)
\label{sec:org1843cbb}
\begin{itemize}
\item it is assumed that we already know the running time at arrival
\item non-preemptive
\item rule only reapplied, when a job is finished (could be also giving back the CPU, while waiting for I/O)
\item better for shorter running jobs, worse for long ones
\end{itemize}
\item Preemptive Shortest Job First
\label{sec:orgfadfd55}
\begin{itemize}
\item it is assumed that we already know the running time at arrival
\item rule reapplied, when a job is finished (+ wait for I/O) or a new job arrives (+ I/O finish)
\item preemptive, resume principle
\item favors the short jobs even more
\item a bit more expensive, because of the context switch
\end{itemize}
\item Round Robin (RR)
\label{sec:org74c80d2}
\begin{itemize}
\item running times not known in advance
\item takes both running and arrival time in consideration
\item cannot make the time quantum too small, because context switch will get too expensive
\item ??? What happens if q = 0???
\item for further info read the heading in OStep/Chapter 7
\end{itemize}
\item Multiple-Level Feedback (FB)
\label{sec:orgb3bef9c}
\begin{itemize}
\item Do not confuse with the modern MLFQ
\item RR but if a task arrives later, it can catch up to the others first
\end{itemize}
\item Two-Level FB / Limited RR
\label{sec:org2737c32}
\begin{itemize}
\item work until a fixed amount of quanta, then put into the background (and only run if no one is in queue 1)
\end{itemize}
\item FB with finite number of levels
\label{sec:org387ecab}
\begin{itemize}
\item just Two-level FB but with a parameter that tells how many queues exist
\item => this gets pretty similar to modern MLFQ
\end{itemize}
\end{enumerate}
\part{Lottery/Stride Scheduling}
\label{sec:org71e0e31}
\part{Adjusting Parameters using Machine Learning}
\label{sec:orgc617b79}
\part{Examples}
\label{sec:orgb85a0df}
\chapter{Linux 2.6 Fair Scheduler}
\label{sec:org0cc4a49}
\chapter{Solaris Scheduling}
\label{sec:orgfa6e5fa}
\end{document}