### 2024-12-31 LaTeX/Thesis.tex

\documentclass{mimosis}

\usepackage{metalogo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Some of my favourite personal adjustments
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% These are the adjustments that I consider necessary for typesetting
% a nice thesis. However, they are *not* included in the template, as
% I do not want to force you to use them.

% This ensures that I am able to typeset bold font in table while still aligning the numbers
% correctly.
\usepackage{etoolbox}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Hyperlinks & bookmarks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[%
  colorlinks = true,
  citecolor  = RoyalBlue,
  linkcolor  = RoyalBlue,
  urlcolor   = RoyalBlue,
  unicode,
  ]{hyperref}

  \usepackage{bookmark}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % Bibliography
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %
  % I like the bibliography to be extremely plain, showing only a numeric
  % identifier and citing everything in simple brackets. The first names,
  % if present, will be initialized. DOIs and URLs will be preserved.

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % Fonts
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \ifxetexorluatex
  \usepackage{unicode-math}
  \setmainfont{EB Garamond}
  \setmathfont{Garamond Math}

  % Load some missing symbols from another font.
  \setmathfont{STIX Two Math}[%
    range = {
      \sharp,
      \natural,
      \flat,
      \clubsuit,
      \spadesuit,
      \checkmark
    }
    ]
    \setmonofont[Scale=MatchLowercase]{Source Code Pro}
    \else
    \usepackage[lf]{ebgaramond}
    \usepackage[oldstyle,scale=0.7]{sourcecodepro}
    \singlespacing
    \fi

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Ordinals
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \makeatletter
    \@ifundefined{st}{%
      \newcommand{\st}{\textsuperscript{\textup{st}}\xspace}
      }{}
      \@ifundefined{rd}{%
        \newcommand{\rd}{\textsuperscript{\textup{rd}}\xspace}
        }{}
        \@ifundefined{nd}{%
          \newcommand{\nd}{\textsuperscript{\textup{nd}}\xspace}
          }{}
          \makeatother

          \renewcommand{\th}{\textsuperscript{\textup{th}}\xspace}

          %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          % Incipit
          %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

          \title{{CPU-Scheduling}}
          \subtitle{What waiting in line at the supermarket and your computer have in common}
          \author{Mark Krutzler 4a}

          %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          % Code Blocks
          %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

          \usepackage{minted}

          \usepackage{hyperref}

          % SVG support
          \usepackage{svg}

          \begin{document}

          \frontmatter
          \include{Sources/Title}
          %\include{Sources/Abstract}

          \tableofcontents

          \mainmatter

          \include{Sources/Basics}

          \include{Sources/Advanced}

          % This ensures that the subsequent sections are being included as root
          % items in the bookmark structure of your PDF reader.
          \bookmarksetup{startatroot}
          \backmatter

          \begin{thebibliography}{99}
            \bibitem{ostep}
              Remzi Hussein Arpaci-Dusseau \& Andrea Carol Arpaci-Dusseau (2018) \emph{Operating Systems: Three Easy Pieces}, CreateSpace Independent Publishing Platform.
            \bibitem{coffman68}
              Edward G. Coffman \& Leonard Kleinrock (1968) \emph{Computer scheduling methods and their countermeasures}, Association for Computing Machinery.
            \bibitem{waldspurger95}
              Carl A. Waldspurger (1995) \emph{Lottery and Stride Scheduling: Flexible Proportional-Share Resource Management}, Massachusetts Institute of Technology.
            \bibitem{waldspurger94}
              Carl A. Waldspurger \& William E. Weihl (1994) \emph{Lottery scheduling: Flexible proportional-share resource management}, Massachusetts Institute of Technology.
            \bibitem{solaris}
              Andrea Carol Arpaci-Dusseau (2000) \emph{Multilevel Feedback Queue Scheduling in Solaris}, available at: \url{https://pages.cs.wisc.edu/~remzi/OSTEP/Citations/notes-solaris.pdf}
          \end{thebibliography}

          \section*{Further Sources}
          \begin{itemize}
            \item \url{https://github.com/Pseudomanifold/latex-mimosis}
            \item \url{https://en.wikipedia.org/wiki/Binary_search}
            \item \url{https://www.quora.com/Why-is-look-up-faster-for-a-Binary-Tree-than-a-Linked-List}
            \item \url{https://ceunican.github.io/aos/09.Scheduling_Proportional_Share.pdf}
            \item \url{https://courses.cs.washington.edu/courses/cse451/12au/l11.pdf}
            \item \url{https://en.wikipedia.org/wiki/Earliest_eligible_virtual_deadline_first_scheduling}
            \item \url{https://en.wikipedia.org/wiki/Proportional_share_scheduling}
            \item \url{https://www.youtube.com/watch?v=MkJfuI5_hjc&t=29s}
            \item \url{https://www.youtube.com/watch?v=qvZGUFHWChY&list=PL9xmBV_5YoZNqDI8qfOZgzbqahCUmUEin&index=1&pp=iAQB}
            \item \url{https://www.youtube.com/watch?v=95s3ndZRGbk&list=PL9xmBV_5YoZNqDI8qfOZgzbqahCUmUEin&index=2&pp=iAQB}
            \item \url{https://en.wikipedia.org/wiki/Red%E2%80%93black_tree} 
          \end{itemize}

          \end{document}